Optimal (Multiway) Spatial Joins

### 2024-05-27 msj.tex
yufei

Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$). A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s$ overlaps with $r$, but $r$ covers neither of the two endpoints of $s$. A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.

@@

Let $S$ be a set of segments where either all segments are horizontal or all are vertical. Given a rectangle $r$, we define 
\myeqn{
    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \}. 
    \label{eqn:cross}
}
Let $R$ be a set of rectangles. Given a horizontal segment, we define 
\myeqn{
    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \}. 
    \label{eqn:contained}
}

Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$. Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle. We define 
\myeqn{
    B_\bm{t} &=& \cap_{i=1}^t r_i \label{eqn:B_t}
}
namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Let us introduce:
\myitems{
    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$. In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.

@@

\extraspacing {\bf Problem $\bm{\mathscr{A}}$.} The input involves a set $P$ of 2D points and set $R$ of rectangles. In the {\em detection version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$. Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time, where $n = |P| + |R|$.

@@

In the {\em reporting version} of the problem, the goal is to output, for each point $p \in P$, all the rectangles $r \in R$ containing $p$, provided that at least one such $r$ exists. For example, in Figure~\ref{fig:guard}a, the output is $\set{(p_2: r_1, r_2), (p_3: r_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.

\extraspacing {\bf Problem $\bm{\mathscr{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments, all in 2D space. The goal is to report, for each segment $h \in H$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$. If $h$ does not intersect with any segment in $V$, report nothing for $h$. Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |V|$.

\extraspacing {\bf Problem $\bm{\mathscr{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ satisfying the condition that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that at least one rectangle $r \in R$ covers both the point $(x_1, y)$ and the point $(x, y)$. If the point $p$ exists (i.e., $h$ has at least one left-end covering rectangle in $R$), we should output a tuple $(h, p)$; otherwise, output nothing for $h$. Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.

\extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, the lowest segment in $\cross_H(r)$ (see \eqref{eqn:cross}). If no segment in $H$ crosses $r$, output nothing for $r$.  Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.

@@

\extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles, all in 2D space. The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ --- defined in \eqref{eqn:contained} --- where the rectangles are sorted by their right boundaries. Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$. Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.

>>>>

Given a horizontal segment $h = [x_1, x_2] \times y$, we call a rectangle $r$ a {\em left-end covering rectangle} of $h$, if $r$ contains the left endpoint of $h$ (i.e., $(x_1, y) \in r$). A horizontal/vertical segment $s$ {\em crosses} a rectangle $r$, if $s \cap r \ne \emptyset$ but $r$ covers neither of the two endpoints of $s$. A rectangle $r$ {\em contains} a horizontal/vertical segment $s$, if $r$ covers both endpoints of $s$.

@@

Let $S$ be a set of segments where either all segments are horizontal or all are vertical. Given a rectangle $r$, we define 
\myeqn{
    \cross_S(r) = \{s \in S \mid \text{$s$ crosses $r$} \}
    \label{eqn:cross}
}
namely, the set of segments in $S$ crossing $r$.
Let $R$ be a set of rectangles. Given a horizontal segment, we define 
\myeqn{
    \contained_R(h) = \{r \in R \mid \text{$h$ is contained in $r$} \}
    \label{eqn:contained}
}
namely, the set of rectangles in $R$ containing $h$.

Given a rectangle $r = [x_1, x_2] \times [y_1, y_2]$, we define $\xleft(r) = x_1$, $\xright(r) = x_2$, $\ybot(r) = y_1$, and $\ytop(r) = y_2$. Consider a $k$-tuple $\bm{t} = (r_1, r_2, ..., r_k)$ where $k \ge 2$, and each $\bm{t}[i] = r_i$ ($i \le [k]$) is a rectangle. We define 
\myeqn{
    B_\bm{t} &=& \cap_{i=1}^t r_i \label{eqn:B_t}
}
namely, $B_\bm{t}$ is the intersection of the rectangles in $\bm{t}$ (note: $B_\bm{t}$ is a rectangle itself). Also, define:
\myitems{
    \item $\gleft(\bm{t})$ as the rectangle $r_i$, $i \in [k]$, satisfying $\xleft(r_i)$ $= \xleft(B_\bm{t})$. In case multiple values in $[k]$ fulfill the condition, let $i$ be the smallest of such values.

@@

\extraspacing {\bf Problem $\bm{\mathscr{A}}$.} The input involves a set $P$ of 2D points and set $R$ of rectangles. In the {\em detection version} of Problem $\mathscr{A}$, the goal is to output, for each point $p \in P$, whether it is covered by at least one rectangle in $R$. Figure~\ref{fig:probs}a gives an example where $P = \set{p_1, p_2, p_3}$ and $R = \set{r_1, r_2}$; the output is ``yes'' for $p_2$ and $p_3$ and ``no'' for $p_1$. This can be achieved in $O(n \log n)$ time where $n = |P| + |R|$ as shown
in Appendix~\ref{app:bricks}.

@@

In the {\em reporting version} of the problem, the goal is to output, for each point $p \in P$, all the rectangles $r \in R$ containing $p$, provided that at least one such $r$ exists. For example, in Figure~\ref{fig:guard}a, the output is $\set{(p_2: r_1, r_2), (p_3: r_2)}$. As shown in Appendix~\ref{app:bricks}, this can be achieved in $O(n \log n + \out)$ time, where $\out$ is the number of pairs $(p, r) \in P \times R$ such that $p \in r$.

\extraspacing {\bf Problem $\bm{\mathscr{B}}$.} The input involves a set $H$ of horizontal segments and a set $V$ of vertical segments. The goal is to report, for each segment $h \in H$, the leftmost point $p$ on $h$ such that $p$ is on some vertical segment in $V$. If $h$ does not intersect with any segment in $V$, report nothing for $h$. Figure~\ref{fig:probs}b gives an example where $H = \set{h_1, h_2, h_3}$ and $V = \set{v_1, v_2}$; the output is $\set{(h_1, p_1), (h_2, p_2)}$. This can be achieved in $O(n \log n)$ time where $n = |H| + |V|$, as shown in Appendix~\ref{app:bricks}.

\extraspacing {\bf Problem $\bm{\mathscr{C}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles. The goal is to report, for each segment $h \in H$, the rightmost point $p$ on $h$ such that $p$ is covered by at least one left-end covering rectangle of $h$ in $R$ --- formally, for $h = [x_1, x_2] \times y$, we aim to find the maximum $x \in [x_1, x_2]$ such that at least one rectangle $r \in R$ covers both the point $(x_1, y)$ and the point $(x, y)$. If the point $p$ exists (i.e., $h$ has at least one left-end covering rectangle in $R$), we should output a tuple $(h, p)$; otherwise, output nothing for $h$. Figure~\ref{fig:probs}c gives an example where $H = \set{h_1, h_2}$ and $R$ includes the three rectangles shown; the output is $\set{(h_1, p_1), (h_2, p_2)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n)$ time where $n = |H| + |R|$.

\extraspacing {\bf Problem $\bm{\mathscr{D}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles. In the {\em find-lowest} version of the problem, the goal is to report, for each rectangle $r \in R$, the lowest segment in $\cross_H(r)$ (see \eqref{eqn:cross}). If no segment in $H$ crosses $r$, output nothing for $r$.  Figure~\ref{fig:probs}d gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2}$; the output is $\set{(r_1, h_3), (r_2, h_2)}$. This can be achieved in $O(n \log n)$ time where $n = |H| + |R|$ as shown
In Appendix~\ref{app:bricks}.

@@

\extraspacing {\bf Problem $\bm{\mathscr{E}}$.} The input involves a set $H$ of horizontal segments and a set $R$ of rectangles. The goal is to report, for each segment $h \in H$, the set $\contained_R(h)$ --- defined in \eqref{eqn:contained} --- where the rectangles are sorted by their right boundaries. Formally, if $r_1, r_2, ..., r_z$ (for some $z \ge 1$) are all the rectangles in $\contained_R(h)$, we output $(h: r_1, r_2, ..., r_z)$, provided that $\xright(r_i) \ge \xright(r_{i-1})$ for each $i \in [2, z]$. Figure~\ref{fig:probs}e gives an example where $H = \set{h_1, h_2, h_3}$ and $R = \set{r_1, r_2, r_3}$; the output is $\set{(h_1: r_2, r_1), (h_2: r_2, r_3)}$. In Appendix~\ref{app:bricks}, we explain how to solve the problem in $O(n \log n + \out)$ time where $n = |H| + |R|$ and $\out$ is the number of pairs $(h, r) \in H \times R$ such that $r$ contains $h$.
